A Switching Planner for Combined Task and Observation Planning

### 2010-11-30 subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/prop-dtp.tex

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process state
$\state$ as the set of propositions that are true of the
state. Notationally, $\props$ is the set of propositions, $\prop$ is
an element from that set, and we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
We say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$.

@@

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for our presentation to clearly
separate sensing and action. Therefore, we suppose a POMDP has a
perceptual model given in terms of a finite set of stochastic {\em
senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
and perceptual propositions $\percepts$, called {\em percepts}. In
detail, we take an observation $\observ$ to be a set of percepts
$\observ \subseteq \percepts$, and denote \observations\ the set of
reachable observations. The underlying state of the process cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in
\observations$ that informs what should be believed about the state a
process is in. If $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from the chosen elements of active
senses.

@@

Until now we have discussed the POMDP in terms of propositions and
percepts. In order to address belief revision and utility it is
convenient to consider the underlying decision process in a flat
format. This is given by the tuple
$\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
The function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state, \action)$. We denote $\obsDist_\observ(\state,
\action)$ the probability of getting observation $\observ$ in state
$\state$. For $\state$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state, \action) = 1$.

Successive state estimation is by application of Bayes'
rule.  Taking the current belief $\bstate$ as the {\em prior}, and
supposing action $\action$ is executed with perceptive outcome
$\observ$, the probability that we are in $\state$ in the
successive belief-state $\bstate'$ is:

\begin{equation}\label{eq:revision}
\bstate'(\state) = \frac{\obsDist_\observ(\state, \action)
  \sum_{\state'\in \states} \transProb(\state', \action, \state) \bstate(\state') }{\pp{Pr}(\observ | \action, \bstate)}
\end{equation}

>>>>

We describe the partially observable propositional probabilistic
planning problem, with costs and rewards. We model a process state
$\state$ as the set of propositions that are true of the
state. Notationally, $\props$ is the set of propositions, $\prop$ is
an element from that set, and we have $\state \subseteq
\props$. The underlying process
dynamics are modelled in terms of a finite set of {\em probabilistic
STRIPS operators}~\cite{boutilier:abstraction} $\stochActions$ over
state-characterising propositions $\propositions$.
%%
We say an action $\stochAction \in \stochActions$ is
applicable if its precondition $\poss(\stochAction)$, a set of
propositions, are satisfied in the current state -- i.e., $\poss(\stochAction) \subseteq
\state$. We denote by $\mu_{\stochAction}(\detAction_i)$ the probability that
nature chooses a deterministic STRIPS effect $\detAct_i$, and for
all \stochAction\ we require
$\sum_{\detAction_i \in \detActions(\stochAction)}
\mu_{\stochAction}(\detAction_i) = 1$.

@@

We are concerned with problems that feature partial
observability. Although we could invoke {\em extended probabilistic
STRIPS operators}~\cite{rintanen:01} to model actions and observations
propositionally, we find it convenient for presentation and computation to
separate sensing and action. Therefore, we suppose a POMDP has a
perceptual model given in terms of a finite set of stochastic {\em
senses} $\stochSenses$, deterministic sensing outcomes $\detSenses$,
and perceptual propositions $\percepts$, called {\em percepts}. In
detail, we take an observation $\observ$ to be a set of percepts
$\observ \subseteq \percepts$, and denote by \observations\ the set of
reachable observations. The underlying state of the process cannot be
observed directly, rather, senses $\stochSense \in \stochSenses$
effect an observation $\observ \in
\observations$ that informs what should be believed about the state the
process is in. If $\action$ is applied effecting a
transition to a successor state $\state'$, then an observation occurs
according to the active senses $\stochSenses(\stochAction, \state')
\subseteq \stochSenses$. A sense $\stochSense$ is active, written
$\stochSense \in \stochSenses(\stochAction, \state')$, if the senses'
action-precondition, $\poss_\stochActions(\stochSense)$, is equal to
$\stochAction$, and the state-precondition $\poss_\states(\stochSense)
\subseteq \propositions$ is satisfied by the state $\state'$ -- In the
usual sense that \mbox{$\poss_\states(\stochSense)\subseteq\state'$}.
%%
When a sense is active, nature must choose exactly one outcome amongst
a small set of deterministic choices $\detSenses(\stochSense)
\equiv \{\detSense_1, \ldots, \detSense_k \}$, so that for each
$i$ we have $\detSense_i \subseteq \percepts$. The probability of
the $i^{th}$ element being chosen is given by
$\psi_{\stochSense}(\detSense_i)$, where $\sum_{\detSense_i \in
\detSenses(\stochSense)} \psi_{\stochSense}(\detSense_i) =
1$. The observation received by the agent corresponds to the union of
perceptual propositions from the chosen elements of active
senses.

@@

Until now we have discussed the POMDP in terms of propositions and
percepts. In order to address belief revision and utility it is
convenient to consider the underlying decision process in a flat
format. This is given by the tuple
$\langle \states, \bstate_0, \actions, \transProb, \reward,
\observations, \obsDist \rangle$. Here, $\bstate_0$ is the initial
belief-state, \states\ is the finite set of reachable propositional
states, \actions\ is the finite set of actions, and \observations\ is
the finite set of reachable observations (i.e., perceptual states).
Where $\state, \state' \in \states$, $a \in \actions$, from $\mu$ we
have a state transition function $\transProb(\state, \action,
\state')$ giving the probability of a transition from state $s$ to
$s'$ if $a$ is applied. For any $\state$ and $\action$ we have
$\sum_{\state' \in \states} \transProb(\state, \action, \state') = 1$.
%%
The function $\reward:\states \times \actions \to \Re$ is a bounded real
valued reward function. Therefore a finite positive constant $c$
exists so that for all $\state \in \states$ and $\action \in
\actions$, $|\reward(\state, \action)| < c$. We model costs as
negative rewards.
%%
From $\psi$ we have that for each $\state' \in \states$ and action
$\action \in \actions$, an observation $\observ \in \observations$ is
generated independently according to a probability distribution
$\obsDist(\state', \action)$. We denote by $\obsDist_\observ(\state',
\action)$ the probability of getting observation $\observ$ in state
$\state'$. For $\state'$ and $\action$ we have $\sum_{\observ \in
\observations} \obsDist_\observ(\state', \action) = 1$.

Successive state estimation is by application of Bayes'
rule.  Taking the current belief $\bstate$ as the {\em prior}, and
supposing action $\action$ is executed with perceptive outcome
$\observ$, the probability that we are in $\state'$ in the
successive belief-state $\bstate'$ is:

\begin{equation}\label{eq:revision}
\bstate'(\state') = \frac{\obsDist_\observ(\state', \action)
  \sum_{\state\in \states} \transProb(\state, \action, \state') \bstate(\state) }{\pp{Pr}(\observ | \action, \bstate)}
\end{equation}
